Reaction Barrier Analysis

### 2024-11-30 Code/ReactionBarrierAnalysis.Rmd
minor
added tukey hsd

---
title: "ReactionBarrierAnalysis"
author: "Owen Fiore"
date: "2024-11-28"
output: html_document
---

The objective of this document is to allow other researchers to replicate our
results in regards to the GAMLSS section of the paper.
# Main Analysis
The first thing we need is our data.  We create the `times` data to be the full
data, and will create relevant subsets depending on what we need to fit.

@@

#Only difference between these two dataframes is the filter removing 2022
dhm_no2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime > 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022) |>
  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
```
Let's visualize the data.
```{r}
boxplot_df <- dhm |>
  mutate(Venue = as.numeric(as.character(Venue)))
library(ggplot2)
ggplot(boxplot_df, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.05, 0.275),
                            breaks = seq(0.05, 0.275, 0.025)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
```
We see that the 2022 boxplot is lower than any other year.  What is particularly
interesting is that Allen's 0.099 reaction time that led to his disqualification
does not appear as an outlier in 2022, it is just the minimum value.  Compared
to in 1999, where a value right around 0.100 was an outlier.

>>>>

---
title: "Reaction Barrier Analysis"
author: "Owen Fiore"
date: "2024-11-28"
output: html_document
---

The objective of this document is to allow other researchers to replicate our
results in regards to the GAMLSS section of the paper.

# Main Analysis

The first thing we need is our data.  We create the `times` data to be the full
data, and will create relevant subsets depending on what we need to fit.

@@

#Only difference between these two dataframes is the filter removing 2022
dhm_no2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime > 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022) |>
  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
```

Let's make sure we used appropriate data. There are three events in the data
file for men: 100 meter dash, 110 meter hurdles, and 200 meter dash.  We
suspect the results from the 200 meter dash might be slower as that is a longer
event and thus reaction time plays a smaller role. We will use a TukeyHSD to
examine any differences between the three events.
```{r}
library(car)
TukeyHSD(aov(RxnTime ~ Event, data = all_events))
```
we see that there are relatively similar results for the 100 meter dash and
110 meter hurdles as the p-value is 0.57 which is quite large.  But either of
these two events when compared with the 200 meter dash results in a statistically
significant p-value and thus we are not going to include this data in our
analysis.

Now that we have validated the decision to exclude certain data, let us 
visualize it.
```{r}
boxplot_df <- dhm |>
  mutate(Venue = as.numeric(as.character(Venue)))
library(ggplot2)
ggplot(boxplot_df, aes(x = Venue, y = RxnTime, group = Venue)) +
  geom_boxplot() +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.05, 0.275),
                            breaks = seq(0.05, 0.275, 0.025)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1999, 2024, 2))
```
We see that the 2022 boxplot is lower than any other year.  What is particularly
interesting is that Allen's 0.099 reaction time that led to his disqualification
does not appear as an outlier in 2022, it is just the minimum value.  Compared
to in 1999, where a value right around 0.100 was an outlier.
